---
title: "Project 2 - Dataset # 3"
author: "J John"
date: "Oct 10, 2018"
output: html_document
---
I am using "Adult Arrests in NYS Beginning in 1970" dataset listed in  discussion board. 


#### Load data
```{r message = FALSE}
require(tidyverse)
 
```

### Arrest Trend
  
 
    Arrests in NY went up from 1970 to late 1990's and peaked around 2010. Arrests have been going down since 2010. 
    

```{r read_file}
arrest_data <- read.csv('ny_adults_arrests.csv')

#Select the Year and Total
total_arrests_by_year <- arrest_data %>% select(Year, Total)  %>% 
                        group_by( Year )  %>% 
                        group_by( Year, total_arrests = sum(Total)) 

ggplot(total_arrests_by_year, aes(x= Year, y= total_arrests)) + geom_line()

arrest_data_long <- arrest_data %>% 
      select(-contains("Total"))  %>%   
      gather(3:10, key = "Type", value = "Total")

arrest_category_year <- arrest_data_long  %>% 
  mutate(Type = str_extract(Type, "[^_]+$") ) %>% 
  group_by(Year, Type)%>% 
  group_by(Year, Type, total_arrests = sum(Total))
```


#### Felony vs Misdemeanor
  
  Same pattern as above is happening for felony and misdemeanor arrests.
  
  
```{r}

ggplot(data = arrest_category_year , aes(x = Year, y= total_arrests, color = Type)) + geom_line()


```


#### Arrests by Type

  This shows arrests further broken down by types. 

```{r arrests_by_type ,eval = TRUE}

total_arrests_by_year_type <-  arrest_data_long %>% 
   select(Year, Type, Total) %>%
  group_by(Year,Type)  %>%
  group_by(Year,  Type, total_arrests = sum(Total)) 

# arrest further broken down by type
ggplot(total_arrests_by_year_type, aes(x= Year, y= total_arrests, color=Type)) + geom_line()
```


#### Arrests by Type
  
  65 % of arrests are Misdemeanor and 35% are felony.

```{r arrest_type}

arrest_category <- arrest_data_long  %>% 
  mutate(Type = str_extract(Type, "[^_]+$") ) %>% 
  group_by( Type)%>% 
  group_by( Type, total_arrests = sum(Total))
 
arrest_num_summary <- summarise(arrest_category)

#% of felony vs misd
round((arrest_num_summary$total_arrests/ sum(arrest_num_summary$total_arrests) ) *100)

ggplot(data = arrest_category , aes(x = Type, y= total_arrests, fill=Type)) +
   geom_bar(stat="identity", position=position_dodge())

total_arrests_by_type <-  arrest_data_long %>%
          select( Type, Total) %>% 
          group_by(Type) %>% 
          group_by(Type, total_arrests = sum(Total)) 

ggplot(total_arrests_by_type, aes(x= Type, y= total_arrests, fill = Type)) + 
  geom_bar(stat="identity", position=position_dodge())

```

### Arrests by County
   

   This shows the number arrests by county. Number of arrests in NYC and surronding counties are higher than upstate NY. Better analaysis requires county population data.
   
   
```{r s}


total_county <- arrest_data_long %>%
          select( County, Total) %>% 
          group_by(County) %>% 
          group_by(County, total_arrests = sum(Total)) 
ggplot(total_county, aes(x= total_arrests, y= County)) + geom_point(shape=1)

```

Arrests in NY peaked around 2000.  Number of arrests have been going down since 2010. 
35% of arrests are felony and 65% are misdemeanor.
"DWI  felony" and  "other felony" arrests aren't down since 2010.
Number of arrests in New York city and surrounding counties are higher than other counties. If we were look at arrests per 1000 people or arrests as percentage of people, I am sure that will make the numbers a lot closer.
