---
title: "Mongo DB"
author: "J John"
date: "Nov 11, 2018"
output:
  
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE, echo= FALSE }
require(tidyr)
require(dplyr)
require(mongolite)
library(RMySQL)
```

#### Mongo Db.
I am using my sql database created  for project #3 and bulk inserting records to a new mongo db.
I wanted to use neo4j for this, but I ran into problems with  neo4j  R package.

### Read Data from project 3 database
```{r }
con = dbConnect(MySQL(), user='data607p3', password='data607p3', dbname='job_skills', host='localhost')


jobs_db <- dbReadTable(con, 'jobs')
skills_db<- dbReadTable(con, 'skills')
job_skills_db<- dbReadTable(con, 'job_skills')

# Join tables to get skills for jobs. 
job_skills_db <- jobs_db %>%
  inner_join(job_skills_db )%>%
  inner_join(skills_db ) 

nrow(job_skills_db)
```
### Create new collection called DATA607 and insert data.
```{r}
data607 <- mongo("DATA607")
data607$insert(job_skills_db)

```
### Number of records inserted.
```{r}
# number of documents 
data607$count()
# number of rows from my sql db
nrow(job_skills_db)
```
### Select everything where skill name is r
```{r}

rskills <- data607$find('{"skill_name" : "r"}')
head(rskills)

```

### Select 50 job names and skill names where job name is 'data scientest'.
```{r}

test <- data607$find(
  query = '{ "job_name" : "data scientist" }', 
  fields = '{"job_name" : true, "skill_name" : true}',
  limit = 50
)

```


Below are some of the pros and cons of sql and no sql databases. 

###SQL
    Predefined schemas and all of your data must follow the same structure. 
    changing  structure would be both difficult and disruptive.
    vertically scalable.
    Good for highly transactional systems.

###NoSQL
    Dynamic schema:  flexibility to change your data schema without modifying any of your existing data.
    Scalability: MongoDB is horizontally scalable.
    Flexibility: You can add new columns or fields on MongoDB without affecting existing rows or application          performance.
    Each document can have its own unique structure.
    No expensive joins