---
title: "File Formats"
author: "J John"
date: "Oct 13, 2018"
output:
  
  html_document:
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE, echo= FALSE}
library(XML)
library(rvest)
library(jsonlite)
library(plyr)
```

#### XML file
Read xml file from github.In this file mutliple authors are listed as coma separated. 

```{r }
xml_url <-"https://raw.githubusercontent.com/jjohn81/Data607/master/Week6/Books.xml"
xml_data <- read_xml(xml_url) %>% xmlParse()
xml_data
xml_df <- xmlToDataFrame(xml_data)
xml_df
```

#### XML file
Read xml file from github. In this file authors are listed as elements.'xmlToDataFrame' cant parse this xml.

```{r }
xml_url <-"https://raw.githubusercontent.com/jjohn81/Data607/master/Week6/Books_muli_authors_element.xml"


xml_data <- read_xml(xml_url) %>% xmlParse()
xml_data
#using plyr moethod. This method creates two columns for authors. 
xml_df_plyr <- ldply(xmlToList(xml_data), data.frame)
xml_df_plyr
```

#### XML file
Read xml file from github. In this file authors are listed in "Authors" node. Dataframe structure depneds on the method used. 
```{r }
xml_url <-"https://raw.githubusercontent.com/jjohn81/Data607/master/Week6/Books_authors_node.xml"
xml_data <- read_xml(xml_url) %>% xmlParse()
xml_data
#Multiple authors are listed under "Authors" column.
xml_df <- xmlToDataFrame(xml_data)
xml_df
#This creates two columns for authors, Authors.author and Authors.author1
xml_df <- ldply(xmlToList(xml_data), data.frame)
xml_df
```

#### Html Table
 Read html file from github.

```{r html}

html_url <- "https://raw.githubusercontent.com/jjohn81/Data607/master/Week6/Books.html"
html_data <- read_html(html_url) %>% html_node("table")
html_data
html_df <-as.data.frame( html_table(html_data,fill=TRUE))
html_df

```

#### JSON 
 Read JSON file from github. In this file mutliple authors are listed as coma separated. 
```{r json}
json_url <-"https://raw.githubusercontent.com/jjohn81/Data607/master/Week6/Books.json"
json_data <- read_json(json_url)
json_data
json_df <-fromJSON(json_url)
json_df

```
 Read JSON file from github. In this file mutliple authors. 
```{r }
json_url <-"https://raw.githubusercontent.com/jjohn81/Data607/master/Week6/books_authors.json"
json_data <- read_json(json_url)
json_data
# creates two author columns. 
json_df <-fromJSON(json_url)
json_df

```

HTML tables are for displaying data while XML and JSON formats are used for storing and transporting data. JSON is lightweight and easy to parse whereas XML is verbose and difficult to parse. Data frame structure depends on the parser/method used. For example, nested authors node might be stored as a data frame column with comma separated names/names concatenated or as multiple columns .