---
title: "Project 1"
author: "J John"
date: "Sep 20, 2018"
output: html_document
---

## Load Package
1. Read all lines. 
2. Remove lines arent required, like --- lines.
3. Split the lines into two lists. One for the rounds info and another with pregame and state     info.
4. Using regular expression parse all required items and store them in indiviudal lists. All 
   individual lists must have 64 items, one per player. I am showing length of the lists to       make sure all lists are of the same size. 
5. Use the iniviudal lists to create a data frame and write that to the file. 

```{r}
require(stringr)
```

## Load data and clean up

```{r read_file}

#Load data from the file using readlines. 
raw_file_lines <- readLines('tournamentinfo.txt', warn = F)

#Remove anything that are not letters and numbers,  like "----"
file_lines <- subset(raw_file_lines, str_detect(raw_file_lines,'[[:alnum:]]'))

```

## Data Summary 
 
```{r file_summary}

#Summary of the original file. Orginal file contains 196 lines including "---"lines.
summary(raw_file_lines)


#Summary of the lines after removing "----" lines
summary(file_lines)

```

## Parse lines
```{r parse_lines, echo= TRUE}
#Split and store the lines into two different lists statrting at line 4 and 3. 
#Player row contains rounds information
player_row <-str_split(file_lines[seq(3, length(file_lines), by=2)], '[|]')
head(player_row, n=1)

#Ratings row contains all pre game player ratings and state
ratings_row <-str_split(file_lines[seq(4, length(file_lines), by=2)], '[|]')
head(ratings_row, n = 1)

#Both lists are of the same size.
length(ratings_row)
length(player_row)
```

##Player names
```{r palyer_names, echo= TRUE}
#Get the player names from the second column.
names <-unlist(lapply(player_row, function(x)str_trim(x[2])))
head(names)
# Number of players
length(names)
```

## Player total points
```{r palyer_points, echo= TRUE}
# Get the total points for the player. 
player_points <- as.numeric(unlist(lapply(player_row, function(x)x[3])))

head(player_points)
length(player_points)
```

## Player states
```{r palyer_states, echo= TRUE}

#Get player state.
state <- unlist(lapply(ratings_row, function(x)str_trim(x[1])))

head(state)
length(state)
```

## Player pre ratings
```{r palyer_ratings, ratings_row, echo= TRUE}

# Second colums contains the pre ratings
ratings_col <-  unlist(lapply(ratings_row, function(x)x[2]))

head(ratings_col)
length(ratings_col)

#Get the pre-rating using regular expression.
player_ratings<-unlist(lapply(ratings_col, function(x) as.numeric(str_extract( str_extract(x,":\\s*\\d{1,}"),"\\d{1,}"))))

head(player_ratings)

length(player_ratings)
```
## Opponents rating
```{r palyer_points,games_row, echo= TRUE}

# Get the rounds player played from column 4 to 10.
player_rounds <- lapply(player_row, function(x)x[c(4:10)])
head(player_rounds, n = 1)
length(player_rounds)

#Get the oppoents Id using regular experession
opponents<-lapply(player_rounds, function(x)unlist(str_extract_all(x,'\\d{1,}')))
head(opponents , n =1)
length(opponents)

#Calculate opponents pre rating.
opponents_ratings <- vapply(opponents, 
                            function(x) round(sum(player_ratings[ as.numeric( unlist(x))])/length(x)),FUN.VALUE = c(0)) 

head(opponents_ratings, n =1)
length(opponents_ratings)
```

## Create a data frame and view summary
```{r create_df, echo= TRUE}
df <-data.frame(names, state,player_points,player_ratings, opponents_ratings)
names(df)<-c('Name','State','Total Points','Pre-Rating','Average Pre-Rating of Opponents')
nrow(df)
head(df)

```

## Write to file
```{r write_to_file, echo= TRUE}
write.table(df,'output.txt',   quote = FALSE, sep = ',', row.names = FALSE)
```

## Read the file 
```{r read_out_put_fiile, echo = TRUE}
data <-read.csv('output.txt')

head(data, n =10)
nrow(data)
```


